Blum axioms and nondeterministic computation of functions

### 2016-04-19 paper.tex
paper.tex: Rewritten section about Krentel's OptP class.

From the last two definitions,
we know we must somehow allow the machine to return several values
and pick only one to be the return value.

We can arrive at the definition proposed by Krentel
analyzing the behavior of a nondeterministic machine to solve the satisfiability problem
that has undergone the process applied to deterministic machines
to transform a decider in a computer for the characteristic function.
More precisely,
pick $M$ a nondeterministic decider for $\SAT$,
and make $M$ write $1$ in the tape if it found a satisfying assignment
and $0$ if it didn't.
The propositional formulas will fall in three categories.
\begin{itemize}
    \item Tautological formulas.
        The set of possible answers for these formulas is only $\{1\}$.
    \item Contradictory formulas.
        Now, the set of answers is $\{0\}$.
    \item Satisfiable but not tautological.
        Here, we have both values: $\{0, 1\}$.
\end{itemize}
The last case is the interesting one.
Note the actual return value for this case is $1$,
which corresponds exactly to the maximum of this set.
We may define the computation using exactly this idea.

\begin{definition}
    If every branch of the computation of $M$ on $x$ halts,
    define the value returned by $M$ on $x$
    as being the lexicographically greatest value
    left by $M$ in any of its branches.
\end{definition}

The requirement that every branch halts
guarantees that this yields an acceptable Gödel numbering;
hence, the first requirement was met.

To apply the Blum axioms,
we define the time complexity by being the size of the deepest branch,
if every branch is finite,
and the space complexity as being the largest number of occupied tape cells
in any branch of computation,
if again every branch is finite.
The requirement to the machine to halt
guarantees we can satisfy both axioms.

Although coming from a different route,
Krentel reached a similar definition \cite{Krentel1988}.
He was concerned with optimization problems
and defined $\OptP$ as being the set of all functions
that could be computed in polynomial time
by nondeterministic Turing machines,
according to the definition of this section.
The idea is that these machines optimize the problem they are solving,
by returning the maximum possible value.
(Actually, Krentel included both maximization and minimization problems
in $\OptP$, but the idea was the same.)

>>>>

The next two authors were not specifically concerned
with nondeterministic computation of functions,
but rather in generalizing the $\NP$ class for functions.
Therefore, the corresponding notion of $\NP$-completeness
behave better than our proposed generalization.

Krentel's definition, in particular,
are very similar to ours.

\begin{definition}
    A function $f: \{0, 1\}^* \to \mathbb N$ is in $\OptP$
    if there is some nondeterministic Turing machine $M$ such that
    \cite[p.~493]{Krentel1988}
    \begin{itemize}
        \item For every input,
            every branch of $M$ halts within a polynomial number of steps
            and writes in its tape a number in binary; and
        \item Either,
            for all $x$, the largest number written by $M$ on $x$ is $f(x)$,
            or, for all $x$, the smallest number written by $M$ on $x$ is $f(x)$.
    \end{itemize}
\end{definition}

Therefore, Krentel's $\OptP$ class contains the optimization problems
that can be ``solved'' in polynomial time by nondeterministic Turing machines.
(Note we must always take the maximum value, or always take the minimum value.
We could associate a Gödel numbering by, say,
adding one bit to the machine encoding.
And, by the same reasoning we made in our definition,
the corresponding complexity measure satisfy Blum axioms.)

The definition of $\OptP$-completeness, however,
is significantly different.

\begin{definition}
    A function $f$ is $\OptP$-complete if $f \in \OptP$ and,
    for every function $g \in \OptP$,
    there are two functions $T_1 : \{0, 1\}^* \to \{0, 1\}^*$
    and $T_2 : \{0, 1\}^* \times \mathbb N \to \{0, 1\}^*$,
    both computable in deterministic polynomial time,
    such that
    \begin{equation*}
        g(x) = T_2( x, f(T_1(x)) ) \qquad \text{for all $x$.}
    \end{equation*}
\end{definition}

That is, besides the preprocessing function $T_1$,
we are allowed to make a post-processing which have access
both to the input $x$ and to the ``reduced output'' $f(T_1(x))$.
Under this definition,
the traveling salesperson problem and $0-1$ integer linear programming
are both $\OptP$-complete \cite[p~495]{Krentel1988}.
